---
title: "Sales Forecasting"
date: "2/5/2021"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

#### Time Series Analysis
|   We will forecast our sales by using weekly data from the data set that we used in our python project. This will give us about 143 observations with which to forecast. First here is a plot of our weekly sales from 2010-02-05 to 2012-10-26.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library("fpp2")
library("tidyverse")
library("lubridate")
library("rmarkdown")
weekly <- read.csv("weekly_sales.csv") %>% select(-X)
weekly_sales <- ts(weekly$Weekly_Sales, start=decimal_date(ymd("2010-02-05")), frequency = 52)
monthly <- read.csv("monthly_sales.csv") %>% 
  select(-X)
monthly_sales <- ts(monthly$Weekly_Sales, start = c(2010), frequency = 12)

monthly_sales.train <- window(monthly_sales, start=c(2010, 2), end=c(2012, 7), frequency=12)
monthly_sales.test <- window(monthly_sales, start=c(2012, 7))

weekly.train <- window(weekly_sales, start=decimal_date(ymd("2010-02-05")), end=decimal_date(ymd("2012-01-06")))
weekly.test <- window(weekly_sales, start=decimal_date(ymd("2012-01-06")))


plot(weekly_sales)
```

|   As you can see our sales seem to spike around the end of the year, most likely because of the holiday sales boom that retail stores typically get. 
|   Before we do any forecasting we need to decide on some baseline models to use as a benchmark. This will allow us to compare seemingly simpler forecasting models with more complicated models. We will use the mean, naive, and seasonal naive as our baseline models. Down below we plot them.

```{r echo=FALSE}

autoplot(weekly.train) +
  autolayer(meanf(weekly.train, h=43),
            series="Mean", PI=FALSE) +
            autolayer(naive(weekly.train, h=43),
                      series="Naive", PI=FALSE) + 
                      autolayer(snaive(weekly.train, h=43),
                                series="Seasonal naive", PI=FALSE) + 
                      ggtitle("Forecasts for Sales Data") + 
                      xlab("Date") + ylab("Sales") + 
                      guides(colour=guide_legend("Forecast"))
```
We first partition our data into train and test splits. The training data is what we use to train our model on. While the testing data is how we evaluate our models performance against data that it hasn't been trained on. Up above the *mean* forecast is merely the historical average of the data. The *naive* forecast is simply the value of the last observation. While the *seasonal naive* forecast is equal to the last observed value from the same period in the previous year.


```{r, echo=FALSE}


```

|   Since our data is highly seasonal we are most concerned with the seasonal naive method. Below we show the accuracy metrics for the seasonal naive method. 

```{r, echo=T, results="hide", include=FALSE}
naive1 <- snaive(weekly.train)
forecast.naive <- forecast(naive1, h=43) 
weekly_sale_summ <- summary(forecast.naive)
weekly_sale_summ <- cbind(Date = rownames(weekly_sale_summ), weekly_sale_summ)
rownames(weekly_sale_summ) <- 1:nrow(weekly_sale_summ)

weekly_sale_summ$Date <- as.numeric(weekly_sale_summ$Date)
weekly_sale_summ$Date <- format(date_decimal(weekly_sale_summ$Date), "%Y-%m-%d")

```

```{r, echo=FALSE}
# forecast.naive %>% accuracy(weekly.test)
# summary(forecast.naive)

accuracy(forecast.naive,weekly.test)


```
As you can see our mean absolute error (MAE) is 1,387,984 for the training set and 1,483,137 for the test set. This means that on average with our testing data we are off by about 1,483,137 from the true value. There's a couple of things to keep in mind when hearing this forecast. First, we only have about 2 years or 104 weeks of weekly data to forecast off of. Ideally we would want several years worth of data to give effective accurate forecasts. This is to ensure that we have enough data to capture and deal with periodic effects such as seasonality. Second, we are forecasting quite far out into the future (about 43 weeks). If we were forecasting not so far out say only a couple of weeks our forecasts and thus MAE would be more accurate.

```{r, echo=FALSE}
plot(forecast.naive)

```
The forecast you see above shows the prediction intervals. The 95% prediction interval is shown in light gray while the 80% interval is shown in dark gray. Below I show our forecast numbers which include the date, point estimate, 80% and 95% prediction interval for all 43 weeks of our forecast.

```{r, echo=FALSE}
weekly_sale_summ
```

### STL Decomposition and Exponential Smoothing
Time series decomposition simply means too break up time series date into different components. One such method is called "Seasonal and Trend decomposition using Loess" or STL for short. We can then use this decomposition to help us forecast values by combining it with another forecasting method. We will now forecast our sales data using STL decomposition along with exponential smoothing (ETS).

```{r, echo=FALSE}
weekly.train.stl <- window(weekly_sales, start=decimal_date(ymd("2010-02-10")), end=decimal_date(ymd("2012-02-17")))

weekly.test.stl <- window(weekly_sales, start=decimal_date(ymd("2012-02-17")))


weekly.train.stl %>% stlf(h=43) %>% autoplot()

weekly.train.stl %>% stlf(h=43) %>% accuracy(weekly.test.stl)

```
As you can see we improved upon our previous seasonal naive model. Our mean absolute error using STL and ETS for the test set is 1,185,812.3. We beat our previous error for the seasonal naive model by about 297,324.7. The MASE error on the far right also confirms this. A MASE below one means that the current model is better than the average naive model.

```{r}

```
It appears that our Exponential Smoothing model beat out our previous baseline model. We can now use this model for forecasting any future sales. Below we list our forecasting for the future 43 weeks using our new model. 

```{r,echo=T, results="hide", include=FALSE}
stlf <- weekly.train.stl %>% stlf(h=43)
weekly_sale_stlf_summ <- summary(stlf)
weekly_sale_stlf_summ <- cbind(Date = rownames(weekly_sale_stlf_summ), weekly_sale_stlf_summ)

rownames(weekly_sale_stlf_summ) <- 1:nrow(weekly_sale_stlf_summ)

weekly_sale_stlf_summ$Date <- as.numeric(weekly_sale_stlf_summ$Date)
weekly_sale_stlf_summ$Date <- format(date_decimal(weekly_sale_stlf_summ$Date), "%Y-%m-%d")

```

```{r, echo=FALSE}
weekly_sale_stlf_summ
```

